---
title: "Project3"
author: "Gabrielle Bartomeo, Binish Chandy, Zach Dravis, Burcu Kaniskan, Niteen Kumar, Betsy Rosalen"
date: "March 25, 2018"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    number_section:  true
    theme: cerulean
    highlight:  tango
editor_options: 
  chunk_output_type: inline
---

```{r global_options, echo=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

```{r setup, include=FALSE}
library(knitr)
library(tidyverse)
library(DT)
library(kableExtra)
library(lubridate)
library(janitor)
library(likert)
library(reshape)
library(plotly)
library(scales)
library(grid)
library(gridExtra)
knitr::opts_chunk$set(echo = FALSE)
```

<br>

# Importing data

The data set used in this project is [Kaggle ML and Data Science Survey 2017](https://www.kaggle.com/kaggle/kaggle-survey-2017/feed). The survey was stored in 2 different data sets:a) multiple choice items, b) free-response items. Kaggle stored each data in csv format. We downloaded the multiple choice item survey results in csv format and placed it in our [GitHub repo](https://github.com/betsyrosalen/DATA_607_Project_3)

Importing Multiple Choice data
```{r message=FALSE, warning=FALSE}
linkMC<-"https://raw.githubusercontent.com/betsyrosalen/DATA_607_Project_3/master/project3_master/rawdata/multipleChoiceResponses.csv"
#importing MC items
MC<-read_csv (linkMC)

survey.data <- MC
#lets create a unique ID variable 
MC$id <- seq.int(nrow(MC))
```

```{r message=FALSE, warning=FALSE}
# Ignore this code Importing conversion rates data incase we want to do analyses 

# link_conversion<-"https://raw.githubusercontent.com/betsyrosalen/DATA_607_Project_3/master/project3_master/rawdata/conversionRates.csv"
# #importing MC items
# conversion<-read_csv (link_conversion)
# dim(conversion)
# #lets create a unique ID variable 
# conversion$id <- seq.int(nrow(conversion))
```

<br>

# Research Question

This project will answer the global research question **Which are the most values data science skills?**

<br>

# Data Science Methods

What are the most frequently used data science (DS) methods by those writing code in DS professions? Do those relate to formal educational attainment?

The Kaggle dataset provides multiple different variables to assess what the most valuable data science skills may be.  In the previous section, we examined what data science methods learners are most excited about and working on.  In this section, we'll look at what which data science methods are the most frequently used and if that has any relationship to educational attainment--a potential indicator of if certain methods require advanced academic training.

## Variables and their definition 

The following variables label questions asking survey respondents how often they use each of these data science methods.  Response options were: Rarely, Sometimes, Often, Most of the time

  - WorkMethodsFrequencyA/B
  - WorkMethodsFrequencyAssociationRules
  - WorkMethodsFrequencyBayesian
  - WorkMethodsFrequencyCNNs
  - WorkMethodsFrequencyCollaborativeFiltering
  - WorkMethodsFrequencyCross-Validation
  - WorkMethodsFrequencyDataVisualization
  - WorkMethodsFrequencyDecisionTrees
  - WorkMethodsFrequencyEnsembleMethods
  - WorkMethodsFrequencyEvolutionaryApproaches
  - WorkMethodsFrequencyGANs
  - WorkMethodsFrequencyGBM
  - WorkMethodsFrequencyHMMs
  - WorkMethodsFrequencyKNN
  - WorkMethodsFrequencyLiftAnalysis
  - WorkMethodsFrequencyLogisticRegression
  - WorkMethodsFrequencyMLN
  - WorkMethodsFrequencyNaiveBayes
  - WorkMethodsFrequencyNLP
  - WorkMethodsFrequencyNeuralNetworks
  - WorkMethodsFrequencyPCA
  - WorkMethodsFrequencyPrescriptiveModeling
  - WorkMethodsFrequencyRandomForests
  - WorkMethodsFrequencyRecommenderSystems
  - WorkMethodsFrequencyRNNs
  - WorkMethodsFrequencySegmentation
  - WorkMethodsFrequencySimulation
  - WorkMethodsFrequencySVMs
  - WorkMethodsFrequencyTextAnalysis
  - WorkMethodsFrequencyTimeSeriesAnalysis

The additional variables used for this analysis will include:

  - Formal Education

## Manipulating data

In order to answer the question of which methods are most popular among code writers, several transformations must first be done.  First, we filter the dataset down to only those who were classified as code writers: those that were employed in some capacity working in data science and writing code as part of their job duties.  Additionally, we include only participants who endorsed at least one data science skill on the question, "At work, which data science methods do you use? (Select all that apply)" with variable name :WorkMethodsSelect.

Once filtered, the endorsed data science methods were aggregated and plotted for frequency (see Exploratory Data Analysis below).  The top five most frequent data science methods endorsed were then selected and given a frequency score to represent among those who endorse using them to some extent, how frequently they use that tool.

The final transformation performed on the data was grouping by formal education level attainment and then identifying the most frequently endorsed data science methods for each group.  This can help identify if those writing certain types of code and using certain data analyses are potentially benefitted by pursuing advanced education--a valuable insight for potential data science pupils.

```{r}
#Subset Data
MethodsData <- MC %>%
  select(FormalEducation, contains("WorkMethods"))

#Filter data based on who answered the methods questions, which were only given to coding workers employed in some capacity
MethodsData <- MethodsData[!is.na(MethodsData$WorkMethodsSelect),]

#Most popular techniques that respondents report using at least rarely
MasterString <- MethodsData$WorkMethodsSelect
Options <- str_c(MasterString, collapse = "")
Options <- unlist(str_split(MasterString, pattern = ","))
Options <- as.data.frame(table(Options))
MethodsFrequency <- arrange(Options, desc(Freq))

#Find out when one knows a technique, how frequently they use it
#Rarely, Sometimes, Often, Most of the time
MethodsScored <- MethodsData %>%
  select(-c(WorkMethodsSelect))
MethodsScored[MethodsScored=="Rarely"]<-1
MethodsScored[MethodsScored=="Sometimes"]<-2
MethodsScored[MethodsScored=="Often"]<-3
MethodsScored[MethodsScored=="Most of the time"]<-4

MethodsScored$WorkMethodsFrequencyDataVisualization <- as.integer(MethodsScored$WorkMethodsFrequencyDataVisualization)
AvgFreqDataVis <- sum(MethodsScored$WorkMethodsFrequencyDataVisualization, na.rm = T) / sum(!is.na(MethodsScored$WorkMethodsFrequencyDataVisualization))

MethodsScored$WorkMethodsFrequencyLogisticRegression <- as.integer(MethodsScored$WorkMethodsFrequencyLogisticRegression)
AvgFreqLogRegress <- sum(MethodsScored$WorkMethodsFrequencyLogisticRegression, na.rm = T) / sum(!is.na(MethodsScored$WorkMethodsFrequencyLogisticRegression))

MethodsScored$`WorkMethodsFrequencyCross-Validation` <- as.integer(MethodsScored$`WorkMethodsFrequencyCross-Validation`)
AvgFreqCrossValid <- sum(MethodsScored$`WorkMethodsFrequencyCross-Validation`, na.rm = T) / sum(!is.na(MethodsScored$`WorkMethodsFrequencyCross-Validation`))

MethodsScored$WorkMethodsFrequencyDecisionTrees <- as.integer(MethodsScored$WorkMethodsFrequencyDecisionTrees)
AvgFreqDecisionTrees <- sum(MethodsScored$WorkMethodsFrequencyDecisionTrees, na.rm = T) / sum(!is.na(MethodsScored$WorkMethodsFrequencyDecisionTrees))

MethodsScored$WorkMethodsFrequencyRandomForests <- as.integer(MethodsScored$WorkMethodsFrequencyRandomForests)
AvgFreqRandomForest <- sum(MethodsScored$WorkMethodsFrequencyRandomForests, na.rm = T) / sum(!is.na(MethodsScored$WorkMethodsFrequencyRandomForests))

Top5FreqScore <- data.frame(Method = c("Data Visualization", "Logistic Regression", "Cross Validation", "Decision Trees", "Random Forests"), FrequencyScore = c(AvgFreqDataVis, AvgFreqLogRegress, AvgFreqCrossValid, AvgFreqDecisionTrees, AvgFreqRandomForest))

#Group and create methods frequency for each level of educational attainment
EducationData <- MethodsData %>%
  select(FormalEducation, WorkMethodsSelect) %>%
  filter(FormalEducation != "I prefer not to answer") %>%
  na.omit()

HighSchoolEducation <- EducationData %>%
  filter(FormalEducation == "I did not complete any formal education past high school")

SomePostSecondaryEducation <- EducationData %>%
  filter(FormalEducation == "Some college/university study without earning a bachelor\'s degree")

ProfessionalEducation <- EducationData %>%
  filter(FormalEducation == "Professional degree")

BachelorsEducation <- EducationData %>%
  filter(FormalEducation == "Bachelor's degree")

MastersEducation <- EducationData %>%
  filter(FormalEducation == "Master's degree")

DoctoralEducation <- EducationData %>%
  filter(FormalEducation == "Doctoral degree")

MethodFrequencyFunction <- function(x){
  BigString <- x$WorkMethodsSelect
  Selections <- str_c(BigString, collapse = "")
  Selections <- unlist(str_split(BigString, pattern = ","))
  TempDF <- as.data.frame(table(Selections))
  TempDF <- TempDF %>%
    mutate(RelativeFreq = Freq / sum(Freq))
  return(TempDF)
}

HighSchoolEducation <- MethodFrequencyFunction(HighSchoolEducation)
SomePostSecondaryEducation <- MethodFrequencyFunction(SomePostSecondaryEducation)
ProfessionalEducation <- MethodFrequencyFunction(ProfessionalEducation)
BachelorsEducation <- MethodFrequencyFunction(BachelorsEducation)
MastersEducation <- MethodFrequencyFunction(MastersEducation)
DoctoralEducation <- MethodFrequencyFunction(DoctoralEducation)
```

## Exploratory Data Analysis (EDA)

Following manipulation of the Kaggle data set, we created plots to visualize the aforementioned research questions.  First, here is a look at the frequency with which the following data science methods were endorsed by a total of 7,773 respondents.  Nearly 2/3 of the survey respondents endorsed the first place skill, data visualization.  Over half endorse logistic regression and just shy of half endorse cross-validation and decision trees.

```{r}
kable(MethodsFrequency)
```

The following plot graphically displays the frequency of endorsements for the data science methods asked about.

```{r}
MethodsFrequencyPlot <- ggplot(MethodsFrequency, aes(x = reorder(Options, -Freq), y = Freq, fill = Options)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  ggtitle("Frequency of Enodrsements for DS Methods") + 
  xlab("Number of Endorsements") +
  ylab("Data Science Method") +
  guides(fill = F)

MethodsFrequencyPlot
```

In this plot we show the "Frequency Score" for the Top Five most endorsed data science methods.  It's important to break this down further than endorsement, as the above table and plot only consider which data science methods one uses at all. Just because a method is endorsed, doesn't mean that individuals use it frequently.  It may be a rare but essential method in data science. To get a more fine grained understanding of how commonly one uses a given data science method on the job, the kaggle survey followed up each endorsed method by asking respondents if they use it Rarely, Sometimes, Often, Most of the time.  We converted these to numeric values (Rarely = 1; Sometimes = 2, Often = 3, and Most of the time = 4) in order to graph a score and average the categorical responses.

```{r}
Top5EndorsedFreqScorePlot <- ggplot(Top5FreqScore, aes(x = reorder(Method, -FrequencyScore), y = FrequencyScore, fill = Method)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  ggtitle("Average Frequency Scores for Top 5 Endorsed DS Methods") + 
  xlab("Average Frequency Score") +
  ylab("Data Science Method")

Top5EndorsedFreqScorePlot
```

Of the top five data science methods endorsed, data visualization was the skill indicated to be used the most frequently.

The below plots show the frequency of methods endorsed for each formal education level assessed by Kaggle.

```{r, fig.height = 15, fig.width = 15, fig.align = "center"}
HighSchoolPlot <- ggplot(HighSchoolEducation, aes(x = Selections, y = RelativeFreq, fill = Selections)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  ggtitle("High School Education Methods Usage") +
  guides(fill = F)

SomeSchoolPlot <- ggplot(SomePostSecondaryEducation, aes(x = Selections, y = RelativeFreq, fill = Selections)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  ggtitle("Some Post Secondary Education Methods Usage") +
  guides(fill = F)

ProfessionalPlot <- ggplot(ProfessionalEducation, aes(x = Selections, y = RelativeFreq, fill = Selections)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  ggtitle("Professional Education Methods Usage") +
  guides(fill = F)

BachelorsPlot <- ggplot(BachelorsEducation, aes(x = Selections, y = RelativeFreq, fill = Selections)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  ggtitle("Bachelor's Education Methods Usage") +
  guides(fill = F)

MastersPlot <- ggplot(MastersEducation, aes(x = Selections, y = RelativeFreq, fill = Selections)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  ggtitle("Master's Education Methods Usage") +
  guides(fill = F)

DoctoralPlot <- ggplot(DoctoralEducation, aes(x = Selections, y = RelativeFreq, fill = Selections)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  ggtitle("Doctoral Education Methods Usage") +
  guides(fill = F)

grid.arrange(HighSchoolPlot, SomeSchoolPlot, ProfessionalPlot, BachelorsPlot, MastersPlot, DoctoralPlot, nrow = 2)
```

We see that in the majority of educational attainment brackets, data visualization remains the most frequently endorsed data science method.

The same information is also provided in tabular format:

```{r}
HighSchoolEducation$Degree <- "High School Education"
SomePostSecondaryEducation$Degree <- "Some Post Secondary Education"
ProfessionalEducation$Degree <- "Professional Education"
BachelorsEducation$Degree <- "Bachelor's Education"
MastersEducation$Degree <- "Master's Education"
DoctoralEducation$Degree <- "Doctoral Education"

AllEducations <- rbind(HighSchoolEducation, SomePostSecondaryEducation, ProfessionalEducation, BachelorsEducation, MastersEducation, DoctoralEducation)

Top3EachDegree <- AllEducations %>%
  group_by(Degree) %>%
  arrange(desc(RelativeFreq)) %>%
  slice(1:3)

kable(Top3EachDegree)
```

Answering the research question of which data science skills are the most important can be interpreted and answered in many ways.  One way to explore this deceivingly complex question is to analyze which data science methods are endorsed as being used by code writers on the job.  This analysis did just that, and further explored the Top 5 most endorsed data science methods by seeing how frequently those that endorsed them actually use those methods on the job.

The bottom line of this analysis is to consider data visualization, logistic regression, cross-validation, decision trees, and random forests as not only frequently endorsed methods, but as methods that are not only essential but used in small ways.  It seems like across data science code writers, these methods are popular and then for individual data science code writers, they are used frequently.

The second goal of this analysis was to understand how formal educational attainment relates to data science methods used on the job.  When looking at the plots of each educational level and the table coalescing all of that data, it does not seem like data science methods used by code writers differ given the educational level.  Data visualization remains the most frequently endorsed data science method for the majority of educational groups.  This has important implications for students of data science in understanding that certain popular job functions are not only performed by those with advanced degrees.  This speaks to how crucial skills like data visualization and the other frequently endorsed and commonly used methods are to data science as a whole.

<br>