---
title: "DATA 607 Final Project"
author: "Zach Dravis"
date: "5/14/2018"
output: html_document
---

# Background and Motivation

# Data acquisition
At least two types of data sources
1. data.gov's data set that looks at the costs providers bill for various services: https://data.cms.gov/Medicare-Inpatient/Inpatient-Prospective-Payment-System-IPPS-Provider/97k6-zzx3
```{r}
CostData <- read.csv("https://raw.githubusercontent.com/zachdravis/CUNY-DATA-607/master/Final%20Project/Inpatient_Prospective_Payment_System__IPPS__Provider_Summary_for_the_Top_100_Diagnosis-Related_Groups__DRG__-_FY2011.csv")
```

# Tidying and Transforming
To tidy these data we need to reduce them to get a single observation something that we can work with. To do this, I want to take the average cost of each diagnosis group in each state.
```{r}
library(stringr)
CostData$State <- str_extract(CostData$Hospital.Referral.Region.Description, "^[:upper:]{2}")

#Need to change the "money" data type to numeric
CostData$Average.Covered.Charges <- str_replace(CostData$Average.Covered.Charges, "\\$", "")
CostData$Average.Covered.Charges <- as.numeric(CostData$Average.Covered.Charges)

CostData$Average.Total.Payments <- str_replace(CostData$Average.Total.Payments, "\\$", "")
CostData$Average.Total.Payments <- as.numeric(CostData$Average.Total.Payments)

CostData$Average.Medicare.Payments <- str_replace(CostData$Average.Medicare.Payments, "\\$", "")
CostData$Average.Medicare.Payments <- as.numeric(CostData$Average.Medicare.Payments)
```


```{r}
library(dplyr)
SummaryCostData <- CostData %>%
  group_by(State, DRG.Definition) %>%
  summarise(CollapsedCoveredCharges = mean(Average.Covered.Charges), CollapsedTotalPayments = mean(Average.Total.Payments), CollapsedMedicarePayments = mean(Average.Medicare.Payments), n = n())
```

Now, I'm interested in a few different things for these data.  First, I want to create a variable that shows the gap between the cost billed and what providers were actually paid.
```{r}
SummaryCostData <- SummaryCostData %>%
  mutate(PaymentGap = CollapsedCoveredCharges - CollapsedTotalPayments)
```

What are the diagnoses with the largest Payment Gaps?
```{r}
DiagnosisData <- SummaryCostData %>%
  filter(n > 10) %>% # Get rid of observations (states) that have 10 or fewer examples of a diagnosis code
  group_by(DRG.Definition) %>%
  summarise(MeanPaymentGap = mean(PaymentGap))

DiagnosisData$DRG.Definition <- as.character(DiagnosisData$DRG.Definition)
DiagnosisData$MeanPaymentGap <- as.numeric(DiagnosisData$MeanPaymentGap)

library(ggplot2)
PaymentGap <- ggplot(DiagnosisData, aes(x = DRG.Definition, y = MeanPaymentGap)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  ggtitle("Frequency of Enodrsements for DS Methods") + 
  xlab("Number of Endorsements") +
  ylab("Data Science Method") +
  guides(fill = F)
```
This is really too difficult to make meaning of.  However, we are in luck in that there are only 100 diagnosis codes.  I can group them based on specialty or area.
```{r}
DiagnosisData$DxGroup <- NA

#First look at brain / head conditions
grep("cranial|head|brain|nerv|stroke|ischemia|seizure|DYSEQUILIBRIUM", DiagnosisData$DRG.Definition, ignore.case = T, value = T)
DiagnosisData$DxGroup[grep("cranial|head|brain|nerv|stroke|ischemia|seizure|DYSEQUILIBRIUM", DiagnosisData$DRG.Definition, ignore.case = T, value = F)] <- "Brain"

#Psych conditions
grep("psych", DiagnosisData$DRG.Definition, ignore.case = T, value = T)
DiagnosisData$DxGroup[grep("psych", DiagnosisData$DRG.Definition, ignore.case = T, value = F)] <- "Psych"

#Heart conditions
grep("circ|heart|chest|cardiac|cardio|myocardial|vascular|atherosclerosis|hypertension|syncope", DiagnosisData$DRG.Definition, ignore.case = T, value = T)
DiagnosisData$DxGroup[grep("circ|heart|chest|cardiac|cardio|myocardial|vascular|atherosclerosis|hypertension|syncope", DiagnosisData$DRG.Definition, ignore.case = T, value = F)] <- "Cardiac"

#GI conditions
grep("Bowel|G\\.I\\.|cholecy|gastro|digest|pancreas", DiagnosisData$DRG.Definition, ignore.case = T, value = T)
DiagnosisData$DxGroup[grep("Bowel|G\\.I\\.|cholecy|gastro|digest|pancreas", DiagnosisData$DRG.Definition, ignore.case = T, value = F)] <- "GI"

#Kidney conditions
grep("renal|kidney", DiagnosisData$DRG.Definition, ignore.case = T, value = T)
DiagnosisData$DxGroup[grep("renal|kidney", DiagnosisData$DRG.Definition, ignore.case = T, value = F)] <- "Kidney"

#Lung conditions
grep("respirat|lung|pulmon|bronchitis", DiagnosisData$DRG.Definition, ignore.case = T, value = T)
DiagnosisData$DxGroup[grep("respirat|lung|pulmon|bronchitis", DiagnosisData$DRG.Definition, ignore.case = T, value = F)] <- "Lung"

#Infections
grep("infect|pneumonia|sepsis|CELLULITIS", DiagnosisData$DRG.Definition, ignore.case = T, value = T)
DiagnosisData$DxGroup[grep("infect|pneumonia|sepsis|CELLULITIS", DiagnosisData$DRG.Definition, ignore.case = T, value = F)] <- "Infection"

#Ortho conditions
grep("hip|joint|spin|back", DiagnosisData$DRG.Definition, ignore.case = T, value = T)
DiagnosisData$DxGroup[grep("hip|joint|spin|back", DiagnosisData$DRG.Definition, ignore.case = T, value = F)] <- "Ortho"

#Overdose/alcohol diagnoses
grep("poison|alcohol|drug", DiagnosisData$DRG.Definition, ignore.case = T, value = T)
DiagnosisData$DxGroup[grep("poison|alcohol|drug", DiagnosisData$DRG.Definition, ignore.case = T, value = F)] <- "Drug"

#Metabolic Conditions and Blood
grep("diabetes|red|nutrition|metabolism", DiagnosisData$DRG.Definition, ignore.case = T, value = T)
DiagnosisData$DxGroup[grep("diabetes|red|nutrition|metabolism", DiagnosisData$DRG.Definition, ignore.case = T, value = F)] <- "Metabolic"

DiagnosisData$DxGroup[is.na(DiagnosisData$DxGroup)] <- "Other" #add other to only one
```

Take a peek at the data to make sure it looks ok and then shorten
```{r}
View(arrange(DiagnosisData, DxGroup))

DiagnosisData <- DiagnosisData %>%
  group_by(DxGroup) %>%
  summarise(MeanPaymentGap = mean(MeanPaymentGap))
```

# Visualize
Graphic that describes or validates data
```{r}
ggplot(DiagnosisData, aes(x = reorder(DxGroup, MeanPaymentGap), y = MeanPaymentGap, fill = MeanPaymentGap)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  ggtitle("Payment Gap For Dx Groups") + 
  xlab("Dx Group") +
  ylab("Average Payment Gap") +
  guides(fill = F)
```

#Transform

Let's see if these means are different using an ANOVA on the whole data set.
```{r}
CostData$DxGroup <- NA

#First look at brain / head conditions
CostData$DxGroup[grep("cranial|head|brain|nerv|stroke|ischemia|seizure|DYSEQUILIBRIUM", CostData$DRG.Definition, ignore.case = T, value = F)] <- "Brain"

#Psych conditions
CostData$DxGroup[grep("psych", CostData$DRG.Definition, ignore.case = T, value = F)] <- "Psych"

#Heart conditions
CostData$DxGroup[grep("circ|heart|chest|cardiac|cardio|myocardial|vascular|atherosclerosis|hypertension|syncope", CostData$DRG.Definition, ignore.case = T, value = F)] <- "Cardiac"

#GI conditions
CostData$DxGroup[grep("Bowel|G\\.I\\.|cholecy|gastro|digest|pancreas", CostData$DRG.Definition, ignore.case = T, value = F)] <- "GI"

#Kidney conditions
CostData$DxGroup[grep("renal|kidney", CostData$DRG.Definition, ignore.case = T, value = F)] <- "Kidney"

#Lung conditions
CostData$DxGroup[grep("respirat|lung|pulmon|bronchitis", CostData$DRG.Definition, ignore.case = T, value = F)] <- "Lung"

#Infections
CostData$DxGroup[grep("infect|pneumonia|sepsis|CELLULITIS", CostData$DRG.Definition, ignore.case = T, value = F)] <- "Infection"

#Ortho conditions
CostData$DxGroup[grep("hip|joint|spin|back", CostData$DRG.Definition, ignore.case = T, value = F)] <- "Ortho"

#Overdose/alcohol diagnoses
CostData$DxGroup[grep("poison|alcohol|drug", CostData$DRG.Definition, ignore.case = T, value = F)] <- "Drug"

#Metabolic Conditions and Blood
CostData$DxGroup[grep("diabetes|red|nutrition|metabolism", CostData$DRG.Definition, ignore.case = T, value = F)] <- "Metabolic"

CostData$DxGroup[is.na(CostData$DxGroup)] <- "Other" #add other to only one

## And now calculate the same payments difference variable
CostData <- CostData %>%
  mutate(PaymentGap = Average.Covered.Charges - Average.Total.Payments)

View(head(CostData))
```


# Model
Now I will run a hypothesis test to see if the payment gap is different for each diagnosis group throughout the country. My Null Hypothesis is that there is no statistically significant difference in the mean of average payments of the diagnostic groups and the alternative hypothesis is that there is at least one mean not equal to the others.
```{r}
hist(CostData$PaymentGap[CostData$DxGroup == "Brain"])
hist(CostData$PaymentGap[CostData$DxGroup == "Infection"])
```

Take a look at these two histograms.  Clearly, the data are not normally distributed, meaning an ANOVA is not warranted here.  After some research, I found the Kruskal–Wallis one-way analysis of variance test, a non parametric statistical test that does not require normality.

```{r}
CostData$DxGroup <- as.factor(CostData$DxGroup)
kruskal.test(PaymentGap ~ DxGroup, data = CostData)
```

Given this p-value, we reject the Null Hypothesis and conclude that different diagnosis groups or "fields" have different gaps in what they are paid back.


What about a percentage of cost?
#Transform
So we will create a new variable that represents the percentage of the cost that was not paid back
```{r}
CostData <- CostData %>%
  mutate(PercentUnpaid = PaymentGap / Average.Covered.Charges)
```

# Visualize
```{r}
CostPercentData <- CostData %>%
  group_by(DxGroup) %>%
  summarise(AvgPercentUnpaid = mean(PercentUnpaid))

ggplot(CostPercentData, aes(x = reorder(DxGroup, AvgPercentUnpaid), y = AvgPercentUnpaid, fill = AvgPercentUnpaid)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  ggtitle("Percent Unpaid For Dx Groups") + 
  xlab("Dx Group") +
  ylab("Percent Unpaid") +
  guides(fill = F)
```

# Model
```{r}
CostData$DxGroup <- as.factor(CostData$DxGroup)
kruskal.test(AvgPercentUnpaid ~ DxGroup, data = CostPercentData)
```
When accounting for the total cost billed out by providers within each diagnosis group, the different diagnosis groups are not paid back significantly differently.

# Transform
Now let's look at things at the state level, collapsed across diagnoses.
```{r}
StatePercentUnpaid <- CostData %>%
  group_by(State) %>%
  summarise(MeanPercentUnpaid = mean(PercentUnpaid))
```

# Visualize
```{r}
ggplot(StatePercentUnpaid, aes(x = reorder(State, MeanPercentUnpaid), y = MeanPercentUnpaid, fill = MeanPercentUnpaid)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  ggtitle("Overall Percent Unpaid by State") + 
  xlab("Dx Group") +
  ylab("Percent Unpaid") +
  guides(fill = F) +
  theme(axis.text.x = element_text( size=12),
          axis.text.y = element_text(size=6))
```

# Model
```{r}
StatePercentUnpaid$State <- as.factor(StatePercentUnpaid$State)
kruskal.test(MeanPercentUnpaid ~ State, data = StatePercentUnpaid)
```
Cannot reject Null Hypothesis.

Now let's look at how cost differs by diagnosis group in each state
# Transform
```{r}
SummaryCostData <- CostData %>%
  group_by(State, DRG.Definition) %>%
  summarise(CollapsedCoveredCharges = mean(Average.Covered.Charges), CollapsedTotalPayments = mean(Average.Total.Payments), CollapsedMedicarePayments = mean(Average.Medicare.Payments), n = n())

SummaryCostData$DxGroup <- NA

#First look at brain / head conditions
SummaryCostData$DxGroup[grep("cranial|head|brain|nerv|stroke|ischemia|seizure|DYSEQUILIBRIUM", SummaryCostData$DRG.Definition, ignore.case = T, value = F)] <- "Brain"

#Psych conditions
SummaryCostData$DxGroup[grep("psych", SummaryCostData$DRG.Definition, ignore.case = T, value = F)] <- "Psych"

#Heart conditions
SummaryCostData$DxGroup[grep("circ|heart|chest|cardiac|cardio|myocardial|vascular|atherosclerosis|hypertension|syncope", SummaryCostData$DRG.Definition, ignore.case = T, value = F)] <- "Cardiac"

#GI conditions
SummaryCostData$DxGroup[grep("Bowel|G\\.I\\.|cholecy|gastro|digest|pancreas", SummaryCostData$DRG.Definition, ignore.case = T, value = F)] <- "GI"

#Kidney conditions
SummaryCostData$DxGroup[grep("renal|kidney", SummaryCostData$DRG.Definition, ignore.case = T, value = F)] <- "Kidney"

#Lung conditions
SummaryCostData$DxGroup[grep("respirat|lung|pulmon|bronchitis", SummaryCostData$DRG.Definition, ignore.case = T, value = F)] <- "Lung"

#Infections
SummaryCostData$DxGroup[grep("infect|pneumonia|sepsis|CELLULITIS", SummaryCostData$DRG.Definition, ignore.case = T, value = F)] <- "Infection"

#Ortho conditions
SummaryCostData$DxGroup[grep("hip|joint|spin|back", SummaryCostData$DRG.Definition, ignore.case = T, value = F)] <- "Ortho"

#Overdose/alcohol diagnoses
SummaryCostData$DxGroup[grep("poison|alcohol|drug", SummaryCostData$DRG.Definition, ignore.case = T, value = F)] <- "Drug"

#Metabolic Conditions and Blood
SummaryCostData$DxGroup[grep("diabetes|red|nutrition|metabolism", SummaryCostData$DRG.Definition, ignore.case = T, value = F)] <- "Metabolic"

SummaryCostData$DxGroup[is.na(SummaryCostData$DxGroup)] <- "Other" #add other to only one

SummaryCostData <- SummaryCostData %>%
  mutate(PercentPaid = CollapsedTotalPayments / CollapsedCoveredCharges)
```

# Model
```{r}
StateGroupModel <- lm(PercentUnpaid ~ State + DxGroup, data = CostData)
summary(StateGroupModel)
```
It looks like a linear model taking into account state and diagnosis group can predict 44% of the variance associated with the percentage paid.



# Acquire
How might these data relate to performance data?

https://data.medicare.gov/Hospital-Compare/Hospital-Value-Based-Purchasing-HVBP-Total-Perform/ypbt-wvdk

```{r}
library(jsonlite)
Jsondata <- "https://data.medicare.gov/resource/4uyb-9g5r.json?$select=*&$limit=3000&$$app_token=dchsMqAAMutE4ypXYatHu9VdF"
Jsondata <- jsonlite::fromJSON(Jsondata)
PerformanceData <- Jsondata
```

#Transform
```{r}
PerformanceData$total_performance_score <- as.numeric(PerformanceData$total_performance_score)
TotalScoreData <- PerformanceData %>%
  group_by(state) %>%
  summarise(MeanTotalPerformance = mean(total_performance_score))
```


```{r, eval=F}
library(sp)
library(rgdal)

US <- readOGR("/Users/zachdravis/Documents/GitHub/CUNY-DATA-607/Final Project/gadm36_USA_shp", "gadm36_USA_1")

US@data$StateTwoLetter <- str_extract(US@data$HASC_1, "..$")

US <- US[US@data$StateTwoLetter %in% TotalScoreData$state,]

US@data <- as.data.frame(left_join(TotalScoreData, US@data, by = c("state" = "StateTwoLetter")))

spplot(US, 'MeanTotalPerformance')
US@data
plot(US[US@data$state=='New York',])
```


```{r}
StateUnpaidPerformance <- left_join(TotalScoreData, StatePercentUnpaid, by = c("state" = "State"))
ggplot(StateUnpaidPerformance, aes(x = MeanTotalPerformance, y = MeanPercentUnpaid)) +
  geom_point(stat = "identity") +
  ggtitle("Frequency of Enodrsements for DS Methods") + 
  xlab("Number of Endorsements") +
  ylab("Data Science Method") +
  geom_smooth(method='lm') +
  guides(fill = F)
```

```{r}
LM_StateUnpaidPerformance <- lm(MeanTotalPerformance ~ MeanPercentUnpaid, data = StateUnpaidPerformance)
summary(LM_StateUnpaidPerformance)
```

```{r}
PerformanceData$provider_number <- as.numeric(PerformanceData$provider_number)
CostPerformanceData <- left_join(PerformanceData, CostData, by = c("provider_number" = "Provider.Id"))
LM_CostPerformance <- lm(total_performance_score ~ PercentUnpaid + State, data = CostPerformanceData)
summary(LM_CostPerformance)
```


# Communicate
Graphic that supports conclusions
Analysis that supports conclusions


Include something not learned in class:
examples: “I used ggmap; I created
a decision tree; I ranked the results; I created my presentation slides directly from R; I figured out to use OAuth 2.0…”

Include a note about challenges