---
title: "DATA607 Assignment 7"
author: "Zach Dravis"
date: "5/2/2018"
output: html_document
---
# Task
Write R code, using your packages of choice, to load the information from each of the three sources into
separate R data frames. Are the three data frames identical?

## Load needed libraries
```{r}
rm(list = ls())
library(RCurl)
library(XML)
library(jsonlite)
```

# HTML
```{r}
HTML_Books <- "https://raw.githubusercontent.com/zachdravis/CUNY-DATA-607/master/Assignment%20Week%207/books.html" #Set URL as object
HTML_Books <- getURLContent(HTML_Books) #Get the html content
HTML_Books <- readHTMLTable(HTML_Books) #Read HTML table
HTML_Books <- HTML_Books[[1]] #Remove from List
HTML_Books <- as.data.frame(HTML_Books) #Create data frame
```

# XML
```{r}
XML_Books <- "https://raw.githubusercontent.com/zachdravis/CUNY-DATA-607/master/Assignment%20Week%207/books.xml" #Set URL as object
XML_Books <- getURLContent(XML_Books) #Get the XML content
XML_Books <- xmlToDataFrame(XML_Books) #Convert it to a dataframe
```

# JSON
```{r}
JSON_Books <- "https://raw.githubusercontent.com/zachdravis/CUNY-DATA-607/master/Assignment%20Week%207/books.json"
JSON_Books <- fromJSON(JSON_Books) #Converts to a list
JSON_Books <- JSON_Books[[1]] #Index the list
JSON_Books <- as.data.frame(JSON_Books) #Convert to Data frame
```

# Examine Differences
```{r}
View(HTML_Books)
View(XML_Books)
View(JSON_Books)
```

Looking at these three data frames, they seem very similar. Let's investigate a bit further
```{r}
str(HTML_Books)
str(XML_Books)
str(JSON_Books)
```

It looks like the HTML and XML tables are the same but that the JSON table stores the authors (the comma separated, multiple values) as a list.  The other strings are stored as characters, whereas in the XML and HTML tables they are stored as factors.

```{r}
JSON_Books[1,3]
JSON_Books[1,4]
```

